A Sound Deep Embedding of Arbitrary Normal Modal Logics in Coq

### 2022-09-04 paper/paper.tex
Fix Löb's theorem section

The start of modal logics can be traced back to Lewis and his definition of 
logical systems with strict implication instead of material implication \cite{Lewis1918}.
Based in this work~\cite{Lewis1932}, independently of classical propositional logic,
defined the systems $\mathsf{S1}$--$\mathsf{S5}$, where $\mathsf{S4}$ and $\mathsf{S5}$
are well known normal modal systems. It was, in fact, G{\"o}del that had the idea 
to build a modal logic system as an extension of the classical propositional calculus, 
when constructing a system similar to $\mathsf{S4}$ in \cite{Godel1933}.

@@

Before defining the Kripke semantics, we need to represent the model used to
interpret a modal logic formula. We encode a frame by a record formed by
a type \mintinline{coq}{W: Set} to represent the set of worlds and an
accessibility relation denoted by a type \mintinline{coq}{R: W -> W -> Prop}.
\begin{minted}{coq}
Record Frame: Type := {
  W: Set;
  R: W -> W -> Prop;
}.
\end{minted}
A model is then denoted by another record, formed by a frame and a labeling function,
which assign the truth value of a variable at a given world.
\begin{minted}{coq}
Record Model: Type := {
  F: Frame; 
  v: nat -> (W F) -> Prop; 
}.
\end{minted}
Note that the type of the labeling function \mintinline{coq}{v} ensures that it
refers to a world from the set of possible worlds in the frame
\mintinline{coq}{F}\footnote{We recall that \mintinline{coq}{W F} is Coq's syntax for
accessing field \mintinline{coq}{W} from the record \mintinline{coq}{F}.}. The encoding of the Kripke semantics is done by a Coq function
that implements the relation of Figure~\ref{fig:kripke}. We omit the code
for propositional logics connectives for brevity. The interesting parts are for
variables (constructor \mintinline{coq}{Lit}) which uses the labeling function
\mintinline{coq}{v} from the model \mintinline{coq}{M}. The meaning of
modalities $\Box$ and $\Diamond$ are represented by a quantifying over worlds
using the accessibility relation $R$. 
\begin{minted}{coq}
Fixpoint formula_valuation (M: Model) (w: W (F M))
                           (phi: formula): Prop :=
  match phi with
  | Lit     x
     => v M x w (* should be inhabited, i.e., true *)
  | Box     psi
     => forall w': W (F M), R (F M) w w' ->
                            formula_valuation M w' psi
  | Dia     psi
     => exists w': W (F M), R (F M) w w' /\
                            formula_valuation M w' psi
  (* some code omitted *)
  end.
\end{minted}
In order to ease the task of writing formulas and its semantics properties, we
define Coq notations that mimics the standard way of writing them.

\paragraph{The Hilbert deductive system}
Basically, a Hilbert-style deductive system allows the use of four
possibilities in a proof: 1) the use of a hypothesis; 2) the use of an axiom;
3) the use of the modus ponens rule or 4) the use of the generalization (necessity)
rule. The inductive type \mintinline{coq}{deduction} encodes these possibilities
as its constructors. The \mintinline{coq}{deduction} type is parameterized by
the set of axioms that form a system for some modal logic. In this way, our definition can be
used by several systems alike.
\begin{minted}{coq}
Inductive deduction (A: axiom -> Prop):
                          theory -> formula -> Prop :=
  (* Premise. *)
  | Prem: forall (t: theory)
                 (f: formula)
                 (i: nat),
          (nth_error t i = Some f) -> deduction A t f
  (* Axiom. *)
  | Ax: forall (t: theory)
               (a: axiom)
               (f: formula),
        A a -> instantiate a = f -> deduction A t f
  (* Modus Ponens. *)
  | Mp: forall (t: theory)
               (f g: formula)
               (d1: deduction A t (f -> g))
               (d2: deduction A t f),
        deduction A t g
  (* Generalization. *)
  | Nec: forall (t: theory)
                (f: formula)
                (d1: deduction A t f),
         deduction A t (Box f).
\end{minted}
The constructors \mintinline{coq}{Prem}, \mintinline{coq}{Mp} and \mintinline{coq}{Nec}
represent rules for hypothesis, modus ponens and generalization, respectively.
Before discussing the \mintinline{coq}{Ax}, we need to discuss how axioms are
represented in our library. The inductive type \mintinline{coq}{axiom} specifies
the different axioms supported by our library.
\begin{minted}{coq}
Inductive axiom: Set :=
  | ax1  : formula -> formula -> axiom
  | ax2  : formula -> formula -> formula  -> axiom
  (* some code omitted *)
  | axK  : formula -> formula -> axiom
  | axPos: formula -> formula -> axiom
  | axT  : formula -> axiom
  | axB  : formula -> axiom
  | axK4 : formula -> axiom
  | axD  : formula -> axiom
  | axK5 : formula -> axiom.
\end{minted}
Each constructor of type \mintinline{coq}{axiom} represents an axiom and
its type stores the subformulas which compose the axiom. The type has
17 constructors: the first 10 consists encode axioms for propositional logics
and the last 7 are for different formulas taken as axioms by systems of modal logics.
Axioms are built using the function \mintinline{coq}{instantiate}, which
receives an axiom and builds its corresponding formula by pattern matching as
follows:
\begin{minted}{coq}
Definition instantiate (a: axiom): formula :=
  match a with
  | ax1    p   q
     => [! p -> (q -> p) !]
  | ax2    p   q   r
     => [! (p -> (q -> r)) -> ((p -> q) -> (p -> r)) !]
  | ax3    p   q
     => [! (~ q -> ~ p) -> (p -> q) !]
  | axK    p   q
     => [! [] (p -> q) -> ([] p -> [] q) !]
  | axPos  p   q
     => [! <> (p \/ q) -> (<> p \/ <> q) !]
  (* some code omitted *)
  end.
\end{minted}
Notation \mintinline{coq}{[! p !]} is used to write modal logic formulas using a
syntax closer to mathematical practice using Coq. This is necessary to avoid
parsing ambiguities, since some symbols are also used by Coq's syntax.
Modalities $\Box$ and $\Diamond$ are denoted by \mintinline{coq}{[]}
and \mintinline{coq}{<>}, respectively.

@@

\paragraph{Properties} A key property of the Hilbert deductive system formalized
in our library is its soundness. If a formula $\varphi$ is derived from a
context $\Gamma$, i.e. $\Gamma\vdash\varphi$, then $\varphi$ is entailed from
the set $\Gamma$, i.e. $\Gamma\models\varphi$. The Coq statement of the soundness
result is:

\begin{minted}{coq}
Theorem soundness:
  forall G p, (K ; G |-- p) -> (G ||= p).
\end{minted}

The proof proceeds by induction on the derivation of $\Gamma \vdash\varphi$ and
it has a case for each of the 11 axioms of the Hilbert system for propositional
logics, one for the $K$ axiom and cases for using hypothesis and the rules modus
ponens. In order to modularize the proof, we defined a lemma for each case.
As example, the following is the statement for modus ponens case.
\begin{minted}{coq}
Lemma Modus_Ponens_soundness:
  forall M w p q,
  ((M ' w ||- p) /\
  (M ' w ||- [! p -> q !])) ->
  (M ' w ||- q).
\end{minted}
All these lemmas are direct consequences from the definitions of the Kripke
semantics and entailment.

\paragraph{Frame Correspondence} 
As already mentioned, there are some modal formula schemas that correspond
to properties of the accessibility relation of a modal system, if we take them
as an axiom in the system.  While the proof that a certain property in a frame
implies the correspondent axiom is usually  direct, the inverse often needs
more effort. On the Coq proofs developed for this work,  the former case is
mostly unfoldings, simplifications and applications of premises
while the latter, apart from the necessity and burden of a proof by contraposition, 
requires the construction of a witness valuation function. 
Nevertheless, all Coq's proofs are fairly similar to the respective
``pen and paper'' proofs that  can be found in the literature. 
Our work proves the correspondence for the  following properties in
accessibility relations: reflexivity, transitivity, symmetry,  euclidean,
seriality, functionality, density, convergency, and converse well-foundness.

@@

In this section we describe a case study of the defined modal logic library.
We proved L\"ob theorem in the system K4 with modal fixed points.
The importance of L\"ob's theorem is due to its use in provability
logic~\cite{boolos_1994} to abstract details from encodings in G\"odel's
incompleteness results.

In order to prove L\"ob's theorem proof we need to encode modal
fixed points. Let $\varphi(A)$ be a formula with only one propositional variable
$A$. We define as a modal fixed point of $\varphi$ is a sentence $\Psi$ such that
$\Psi \leftrightarrow \varphi(\Box \Psi)$.
We represent the modal fixed points by definition \mintinline{coq}{fixed_point},
which is parameterized by a system for modal logics, \mintinline{coq}{S}, and a
theory, \mintinline{coq}{g}, and encodes the logical equivalence between a
formula \mintinline{coq}{p} and its fixpoint \mintinline{coq}{f ([] p)}. 
\begin{minted}{coq}
  Definition fixed_point S g
      := forall f, exists p,
             (S ; g |-- [! (p <-> f ([] p)) !]).
\end{minted}

\noindent Using the definition of fixed points, we can state L\"ob's theorem as follows:

\begin{minted}{coq}
  Theorem Lob: forall P, fixed_point K4 nil ->
               (K4; nil |-- [! []P -> P !]) ->
               (K4; nil |-- [! P !]).
\end{minted}

\noindent The proof starts by assuming an arbitrary formula $P$, the existence of modal fixed
points and that $\Box P \to P$ is provable in K4. These deduction steps
are performed by the following tactic:

\begin{minted}{coq}
  intros P FP H1.
\end{minted}

@@

\begin{minted}{coq}
  destruct FP with (fun A => [! A -> P !]) as (psi, H2).
\end{minted}

@@

\begin{minted}{coq}
  assert (K4; nil |-- [! psi -> []psi -> P !]) as H3.
  apply modal_ax5 in H2; try repeat constructor ;
     try eassumption.
\end{minted}

\noindent The \mintinline{coq}{assert} tactic allows the definition of
auxiliary proof goals. In the previous code piece, we specified
$\vdash \Psi \to (\Box \Psi \to P)$ as a new goal which
generates a new proof hypothesis named \mintinline{coq}{H3}. The
deduction of this fact is immediate from K4 axioms.

From $\vdash \Psi \to (\Box \Psi \to P)$ and the necessitation
rule, we can deduce $\vdash \Box (\Psi \to (\Box \Psi \to P))$
using the following tactics:
\begin{minted}{coq}
  assert (K4; nil |-- [! [](psi -> []psi -> P) !])
    as H4.
  apply Nec; auto.
\end{minted}

@@

\begin{minted}{coq}
  assert (K4; nil |-- [! []psi -> []([]psi -> P) !])
    as H5.
  apply modal_axK in H4; try repeat constructor ;
    try eassumption.
\end{minted}

@@

\begin{minted}{coq}
  assert (K4; nil |-- [! []([]psi -> P) ->
                      [][]psi -> []P !]) as H6.
  eapply Ax with (a := axK ?[X] ?[Y]);
     try repeat constructor.
\end{minted}

@@

\begin{minted}{coq}
  assert (K4; nil |-- [! []psi -> [][]psi -> []P !])
    as H7.
  eapply modal_compose; try repeat constructor ;
    try eassumption.
\end{minted}

\noindent Now, we can show that $\vdash\Box\Psi\to\Box\Box\Psi$ using
the axioms of K4. The next snippet of tactics perform this
reasoning step.

\begin{minted}{coq}
  assert (K4; nil |-- [! []psi -> [][]psi !]) as H8.
  apply modal_axK4; try constructor 2.
\end{minted}

\noindent From $\vdash \Box\Psi \to \Box P$ and $\vdash \Box P \to P$, we can
conclude $\vdash \Box \Psi \to P$, by composition of implication as
follows.

\begin{minted}{coq}
  assert (K4; nil |-- [! []psi -> P !]) as H10.
    eapply modal_compose; try repeat constructor ;
      try eassumption.
\end{minted}

\noindent Next, we can conclude $\vdash \Box (\Psi \to P) \to \Psi$ from
$\Psi \leftrightarrow (\Box \Psi \to P)$ and the axioms of K4 using
the following tactics:

\begin{minted}{coq}
  assert (K4; nil |-- [! ([]psi -> P) -> psi !]) as H11.
  apply modal_ax6 in H2; try repeat constructor ;
     try eassumption.
\end{minted}

\noindent From $\vdash (\Box\Psi \to P) \to \Psi$ and $\vdash \Box \Psi \to P$
we can conclude $\vdash \Psi$ by modus ponens.

\begin{minted}{coq}
  assert (K4; nil |-- psi) as H12.
  eapply Mp ; try eassumption.
\end{minted}

@@

\begin{minted}{coq}
  assert (K4; nil |-- [! []psi !]) as H13.
  apply Nec ; try eassumption.
\end{minted}

@@

\begin{minted}{coq}
  eapply Mp ; try eassumption.
\end{minted}

\noindent This concludes the formalization of L\"ob's theorem using
the developed library.

@@

The application of a multimodal separation logic for the semantics of low-level
was the subject of Dockins et. al. work~\cite{Dockins08}. In that work, authors
combine several modal operators that do not need to satisfy the same frame
conditions and seperation logics for reasoning about non-trivial subsets of
low-level programming languages. Authors developed their formalization through a
shallow embedding in Coq proof assistant and provide two case studies:
the verification of a low-level virtual machine and a C-like language.

A model for recursive, impredicatively quantified types with mutable references
was studied by Appel et. al.~\cite{Appel07}. Their interest in formalizing such
advanced typing concepts was in the context of typed assembly languages.
The authors stabilished a semantic type soundness proof by defining a Kripke-style 
semantics of the G\"odel-L\"ob logic of provability. All their definitions were
mechanized through a shallow embedding in Coq.

Finally, given that a deep embedding for the modal logic was used, it's
natural to wonder how that might interact with the program extraction feature. The
reader might refer to~\cite{Kavvos16c}, which presents a survey on the computational
interpretation of several modal logic systems. Whether or not the current implementation
is suitable for extraction is left for future work.

>>>>

The start of modal logics can be traced back to Lewis and his definition of 
logical systems with strict implication instead of material implication \cite{Lewis1918}.
Based in this work, independently of classical propositional logic,
defined the systems $\mathsf{S1}$--$\mathsf{S5}$~\cite{Lewis1932}, where $\mathsf{S4}$ and $\mathsf{S5}$
are well known normal modal systems. It was, in fact, G{\"o}del that had the idea 
to build a modal logic system as an extension of the classical propositional calculus, 
when constructing a system similar to $\mathsf{S4}$ in \cite{Godel1933}.

@@

Before defining the Kripke semantics, we need to represent the model used to
interpret a modal logic formula. We encode a frame by a record formed by
a type \mintinline{coq}{W: Set} to represent the set of worlds and an
accessibility relation denoted by a type \mintinline{coq}{R: W -> W -> Prop}.
\begin{minted}{coq}
Record Frame: Type := {
  W: Set;
  R: W -> W -> Prop;
}.
\end{minted}
A model is then denoted by another record, formed by a frame and a labeling function,
which assign the truth value of a variable at a given world.
\begin{minted}{coq}
Record Model: Type := {
  F: Frame; 
  v: nat -> (W F) -> Prop; 
}.
\end{minted}
Note that the type of the labeling function \mintinline{coq}{v} ensures that it
refers to a world from the set of possible worlds in the frame
\mintinline{coq}{F}\footnote{We recall that \mintinline{coq}{W F} is Coq's syntax for
accessing field \mintinline{coq}{W} from the record \mintinline{coq}{F}.}. The encoding of the Kripke semantics is done by a Coq function
that implements the relation of Figure~\ref{fig:kripke}. We omit the code
for propositional logics connectives for brevity. The interesting parts are for
variables (constructor \mintinline{coq}{Lit}) which uses the labeling function
\mintinline{coq}{v} from the model \mintinline{coq}{M}. The meaning of
modalities $\Box$ and $\Diamond$ are represented by quantifying over worlds
using the accessibility relation $R$. 
\begin{minted}{coq}
Fixpoint formula_valuation (M: Model) (w: W (F M))
                           (phi: formula): Prop :=
  match phi with
  | Lit     x
     => v M x w (* should be inhabited, i.e., true *)
  | Box     psi
     => forall w': W (F M), R (F M) w w' ->
                            formula_valuation M w' psi
  | Dia     psi
     => exists w': W (F M), R (F M) w w' /\
                            formula_valuation M w' psi
  (* some code omitted *)
  end.
\end{minted}
In order to ease the task of writing formulas and its semantics properties, we
define Coq notations that mimics the standard way of writing them.

\paragraph{The Hilbert deductive system}
Basically, a Hilbert-style deductive system allows the use of four
possibilities in a proof: 1) the use of a hypothesis; 2) the use of an axiom;
3) the use of the modus ponens rule or 4) the use of the generalization (necessity)
rule. The inductive type \mintinline{coq}{deduction} encodes these possibilities
as its constructors. The \mintinline{coq}{deduction} type is parameterized by
the set of axioms that form a system for some modal logic. In this way, our definition can be
used by several systems alike.
\begin{minted}{coq}
Inductive deduction (A: axiom -> Prop):
                          theory -> formula -> Prop :=
  (* Premise. *)
  | Prem: forall (t: theory)
                 (f: formula)
                 (i: nat),
          (nth_error t i = Some f) -> deduction A t f
  (* Axiom. *)
  | Ax: forall (t: theory)
               (a: axiom)
               (f: formula),
        A a -> instantiate a = f -> deduction A t f
  (* Modus Ponens. *)
  | Mp: forall (t: theory)
               (f g: formula)
               (d1: deduction A t (f -> g))
               (d2: deduction A t f),
        deduction A t g
  (* Generalization. *)
  | Nec: forall (t: theory)
                (f: formula)
                (d1: deduction A t f),
         deduction A t (Box f).
\end{minted}
The constructors \mintinline{coq}{Prem}, \mintinline{coq}{Mp} and \mintinline{coq}{Nec}
represent rules for hypothesis, modus ponens and generalization, respectively.
Before discussing the \mintinline{coq}{Ax}, we need to discuss how axioms are
represented in our library. The inductive type \mintinline{coq}{axiom} specifies
the different axioms supported by our library.
\begin{minted}{coq}
Inductive axiom: Set :=
  | ax1  : formula -> formula -> axiom
  | ax2  : formula -> formula -> formula  -> axiom
  (* some code omitted *)
  | axK  : formula -> formula -> axiom
  | axPos: formula -> formula -> axiom
  | axT  : formula -> axiom
  | axB  : formula -> axiom
  | axK4 : formula -> axiom
  | axD  : formula -> axiom
  | axK5 : formula -> axiom.
\end{minted}
Each constructor of type \mintinline{coq}{axiom} represents an axiom and
its type stores the subformulas which compose the axiom. The type has
17 constructors: the first 10 encode axioms for propositional logics
and, the last 7 are for different formulas which are taken as axioms by common
systems of modal logic. Axioms are built using the function \mintinline{coq}{instantiate},
which receives an axiom and builds a corresponding formula by pattern matching as
follows:
\begin{minted}{coq}
Definition instantiate (a: axiom): formula :=
  match a with
  | ax1    p   q
     => [! p -> (q -> p) !]
  | ax2    p   q   r
     => [! (p -> (q -> r)) -> ((p -> q) -> (p -> r)) !]
  | ax3    p   q
     => [! (~ q -> ~ p) -> (p -> q) !]
  | axK    p   q
     => [! [] (p -> q) -> ([] p -> [] q) !]
  | axPos  p   q
     => [! <> (p \/ q) -> (<> p \/ <> q) !]
  (* some code omitted *)
  end.
\end{minted}
Notation \mintinline{coq}{[! p !]} is used to write modal logic formulas using a
syntax closer to mathematical practice using Coq. This is necessary to avoid
parsing ambiguities, since some symbols are also used by Coq's syntax.
Modalities $\Box$ and $\Diamond$ are denoted by \mintinline{coq}{[]}
and \mintinline{coq}{<>}, respectively.

@@

\paragraph{Properties} A key property of the Hilbert deductive system as formalized
in our library is its soundness (through system K). If a formula $\varphi$ is derived from a
context $\Gamma$, i.e. $\Gamma\vdash\varphi$, then $\varphi$ is entailed within the deep embedding
from the set of hypotheses $\Gamma$, i.e. $\Gamma\models\varphi$. The Coq statement for the soundness
result is given as follow:

\begin{minted}{coq}
Theorem soundness:
  forall G p, (K; G |-- p) -> (G ||= p).
\end{minted}

The proof proceeds by induction on the derivation of $\Gamma \vdash\varphi$ and
it has a case for each of the 11 axioms of the Hilbert system for propositional
logics, one for the $K$ axiom and cases for using hypothesis and the rules modus
ponens. In order to modularize the proof, we defined a lemma for each case.
As example, the following is the statement for modus ponens case.
\begin{minted}{coq}
Lemma Modus_Ponens_soundness:
  forall M w p q,
  (M ' w ||- p) /\ (M ' w ||- [! p -> q !]) ->
  (M ' w ||- q).
\end{minted}
All these lemmas are direct consequences from the definitions of the Kripke
semantics and entailment. We note that system K is used as it doesn't impose
any restrictions on the frame.

\paragraph{Frame Correspondence} 
As already mentioned, there are some modal formula schemas that correspond
to properties of the accessibility relation of a modal system, if we take them
as an axiom in the system.  While the proof that a certain property in a frame
implies the correspondent axiom is usually  direct, the inverse often needs
more effort. On the Coq proofs developed for this work,  the former case is
mostly unfoldings, simplifications and applications of premises
while the latter, apart from the necessity and burden of a proof by contraposition, 
requires the construction of a witness valuation function. 
Nevertheless, all Coq's proofs are fairly similar to the respective
``pen and paper'' proofs found in the literature. 
Our work proves correspondence for the following properties: reflexivity,
transitivity, symmetry,  euclidean, seriality, functionality, density, convergency,
and converse well-foundness.

@@

In this section we describe a case study of the defined modal logic library:
a proof L\"ob theorem in an extension of system K4 with modal fixed points.
The importance of L\"ob's theorem is due to its use in provability
logic~\cite{boolos_1994}, used to abstract details from encodings in G\"odel's
incompleteness results.

In order to state L\"ob's theorem we need a way to encode modal
fixed points. Formally, let $\varphi(X)$ be a formula with only one propositional variable
$X$. The modal fixed point of $\varphi$ is then a formula $\Psi$ such that such that
$\Psi \leftrightarrow \varphi(\Box \Psi)$.
We represent the existence of modal fixed points for all one-variable formulas by a definition \mintinline{coq}{fixed_point},
which is parameterized by a system of modal logic, \mintinline{coq}{S}, and a
theory, \mintinline{coq}{g}, and encodes the logical equivalence demonstrating
that \mintinline{coq}{p} is a modal fixed point for \mintinline{coq}{f}.
\begin{minted}{coq}
  Definition fixed_point S g: Prop :=
    forall f, exists p,
    (S; g |-- [! (p <-> f ([]p)) !]).
\end{minted}

\noindent We can now write down L\"ob's theorem as follows:

\begin{minted}{coq}
  Theorem Lob:
    forall A, subset K4 A /\ fixed_point A nil ->
    forall P, (A; nil |-- [! []P -> P !]) ->
              (A; nil |-- [! P !]).
\end{minted}

\noindent The proof starts by assuming an arbitrary formula $P$, the existence of modal fixed
points and that $\Box P \to P$ is provable with axioms A. These deduction steps
are performed by the following tactic:

\begin{minted}{coq}
  intros A [I FP] P H1.
\end{minted}

@@

\begin{minted}{coq}
  destruct FP with (fun X => [! X -> P !]) as (psi, H2).
\end{minted}

@@

\begin{minted}{coq}
  assert (A; nil |-- [! psi -> []psi -> P !]) as H3.
  apply modal_ax5 in H2; auto with modal.
\end{minted}

\noindent The \mintinline{coq}{assert} tactic allows the definition of
auxiliary proof goals. In the previous code piece, we specified
$\vdash \Psi \to (\Box \Psi \to P)$ as a new goal which
generates a new proof hypothesis named \mintinline{coq}{H3}. The
deduction of this fact is immediate from K4 axioms, of which A is a superset.
We use some basic automation to move on.

From $\vdash \Psi \to (\Box \Psi \to P)$ and the necessitation
rule, we can deduce $\vdash \Box (\Psi \to (\Box \Psi \to P))$
using the following tactics:
\begin{minted}{coq}
  assert (A; nil |-- [! [](psi -> []psi -> P) !])
    as H4.
  apply Nec; auto.
\end{minted}

@@

\begin{minted}{coq}
  assert (A; nil |-- [! []psi -> []([]psi -> P) !])
    as H5.
  apply modal_axK in H4; auto with modal.
\end{minted}

@@

\begin{minted}{coq}
  assert (A; nil |-- [! []([]psi -> P) ->
                          [][]psi -> []P !]) as H6.
  eapply Ax with (a := axK ?[X] ?[Y]); auto with modal.
  reflexivity.
\end{minted}

@@

\begin{minted}{coq}
  assert (A; nil |-- [! []psi -> [][]psi -> []P !])
    as H7.
  eapply modal_compose; eauto with modal.
\end{minted}

\noindent Now, we can show that $\vdash\Box\Psi\to\Box\Box\Psi$ using
the axiom K4 of A. The next snippet of tactics perform this
reasoning step.

\begin{minted}{coq}
  assert (A; nil |-- [! []psi -> [][]psi !]) as H8.
  apply modal_axK4; auto with modal.
\end{minted}

\noindent From the previou two hypotheses we gather necessary info to derive
$\vdash \Box \Psi \to P$, using a propositional axiom.

\begin{minted}{coq}
  assert (A; nil |-- [! []psi -> []P !]) as H9.
  eapply modal_ax2; eauto with modal.
\end{minted}

\noindent Then, from $\vdash \Box\Psi \to \Box P$ and $\vdash \Box P \to P$, we can
conclude $\vdash \Box \Psi \to P$, by composition of implication as
follows.

\begin{minted}{coq}
  assert (A; nil |-- [! []psi -> P !]) as H10.
  eapply modal_compose; eauto with modal.
\end{minted}

\noindent Next, we can conclude $\vdash \Box (\Psi \to P) \to \Psi$ from
$\Psi \leftrightarrow (\Box \Psi \to P)$ using
the following tactics:

\begin{minted}{coq}
  assert (A; nil |-- [! ([]psi -> P) -> psi !]) as H11.
  apply modal_ax6 in H2; auto with modal.
\end{minted}

\noindent Now, from $\vdash (\Box\Psi \to P) \to \Psi$ and $\vdash \Box \Psi \to P$
we can conclude $\vdash \Psi$ by modus ponens.

\begin{minted}{coq}
  assert (A; nil |-- psi) as H12.
  eapply Mp; try eassumption.
\end{minted}

@@

\begin{minted}{coq}
  assert (A; nil |-- [! []psi !]) as H13.
  apply Nec; try eassumption.
\end{minted}

@@

\begin{minted}{coq}
  eapply Mp; try eassumption.
\end{minted}

\noindent This concludes the formalization of L\"ob's theorem using
the developed library, in a Hilbert style proof.

@@

The application of a multimodal separation logic for the semantics of low-level languages
was the subject of Dockins et. al. work~\cite{Dockins08}. In that work, authors
combine several modal operators that do not need to satisfy the same frame
conditions and seperation logics for reasoning about non-trivial subsets of
low-level programming languages. Authors developed their formalization through a
shallow embedding in Coq proof assistant and provide two case studies:
the verification of a low-level virtual machine and a C-like language.

A model for recursive, impredicatively quantified types with mutable references
was studied by Appel et. al.~\cite{Appel07}. Their interest in formalizing such
advanced typing concepts was in the context of typed assembly languages.
The authors stabilished a semantic type soundness proof by defining a Kripke-style 
semantics of the G\"odel-L\"ob logic of provability. All their definitions were
also mechanized through a shallow embedding in Coq.

Finally, given that a deep embedding for the modal logic was used, it's
natural to wonder how that might interact with the program extraction feature. The
reader might refer to~\cite{Kavvos16c}, which presents a survey on the computational
interpretation of several modal logic systems. Whether or not our current implementation
is suitable for extraction is left for future work.
